---

output: github_document
---
# PROGRAMING ASSIGNMENT 3
```{r} 

library("dplyr")   
library("ggplot2") 
library("readr")   
library("here")    
library("tidyr")
library("stringr")
```
# Answering Questions from Task 7.
1) "vonset = Get starting point:2,2" refers to the start (onset) of the the vowel (as marked on the textgrid) and the number 2 refers to the tier in which this measurement is located (i.e., second tier of the textgrid). The same for the following line, "voffset = Get end point: 2,2", with the exception that this time, it's about the end of the vowel (as marked on the textgrid). The third line of code, "durationV = voffset - vonset" simply calculates the duration of the vowel through a substraction of the lower value (i.e., vonset) from the higher value (i.e., voffset). 

2) The general outline of the script can be summarized as follows:
 - We first set up the output file and tell the divine forces of coding where we want to save the file (i.e, establish a path). At this point, we are also naming the file and naming the headers (e.g., *item*, *vowel*, etc.)
 - Then, things get a bit more interesting and we set up a loop. In order to do this, we first set the path to the stim files, since this is where the divine forces of coding will find the needed information to get values.
 - At this point, we are basically telling the code what to do: go to the wav and textgrid files and check the data (that's where you're going to get all the info from!).
 - Then, things get more interesting and the code is working on *getting the item and vowel label*, as well as the *language label and time landmark*. 
 -Similarly, the code is extracting the values (measurements) that we are asking it to extract for f1 and f2 and tl.
 - Finally, we are telling the code to *print the results to window and save to.csv file*, which is coding lingo for *just save the values that you've been working on calculating in a specific file that has the .csv format*

3) The first difference between the pa2 and pa3 is that for pa2, the textgrid contained only one tier (in which we included only the word and a digit representing the stressed syllable). For this pa3, we have three tiers. 
 For pa2, we recorded all the words in one file and  segmented them, but then we used the code to take the segmented words and create individual files (wav and textgrid) for each one of them. So, we ended up with a *segmented* folder that contained multiple wav and textgrid files (corresponding to the words we recorded and segmented). For pa3, we also recorded multiple words (embedded in carrier sentences) and segmented them but we did not use a code to create another folder that had a wav and a textgrid file for each one of the recordings. 
 Then, for pa2, we had to use another script to extract the values that we wanted for each individual word.
 
 I think there are advantages and disadvantages for both. I think the script we used for pa3 makes more sense if we have longer bits of data, such as interviews in which participants are spontaneously producing data (aka we are not controlling what they say). Let's say we wanted to extract all the "a" vowels that the participant produces. I think the type of script we used for pa3 might make more sense. 
 But, if we were to have different participants produce the same data, a script like the one for pa2 might make more sense, because then we could compare the segmented data across participants (e.g., look at everyone's production of the vowel "a" in *casa*). Ultimately, I'm not sure if these are actual advantages or disadvantages, but these are some of the things I can think of.
 
# Load and read vowel data 
```{r}
here("data")
```
```{r}
read_csv(file = "./data/vowel_data.csv")
```
```{r}
my_data <- read_csv(here("data", "vowel_data.csv"))
```

# Manipulate dataframe to calculate F1/F2 centroids and Trajectory Lengths
```{r}
vowels<-select(my_data,vowel, language,f1_cent,f2_cent,tl)
```
```{r}
vowels_calc <- vowels%>% 
  group_by(vowel,language) %>% 
  summarize(f1_mean = mean(f1_cent), f1_sd = sd(f1_cent),
            f2_mean = mean(f2_cent), f2_sd = sd (f2_cent), 
                           tl_mean = mean (tl), tl_sd = sd(tl))
```

## However, if we wanted to have the average values separated into three distinct categories/groups, such as (1) F1 centroid, (2) F2 centroid, and (3) Trajectory Length, we could also do this:

### F1 mean
```{r}
f1_mean <- vowels%>% 
  group_by(vowel,language) %>% 
  summarize(f1_mean = mean(f1_cent)) %>% 
  pivot_wider(names_from = language,
              values_from = f1_mean)
```
### F2 mean
```{r}
f2_mean <- vowels%>% 
  group_by(vowel,language) %>% 
  summarize(f2_mean = mean(f2_cent)) %>% 
  pivot_wider(names_from = language,
              values_from = f2_mean)
```

### TL mean
```{r}
tl_mean <- vowels%>% 
  group_by(vowel,language) %>% 
  summarize(tl_mean = mean(tl)) %>% 
  pivot_wider(names_from = language,
              values_from = tl_mean)
```

# Plots

## 1. Trajectory Length as a function of vowel and language

```{r}
vowels %>% 
  ggplot(.,aes(x=tl, y=vowel, color=language))+
  geom_point()
```

## 2. F1 as a function of vowel and language

```{r}
vowels %>% 
  ggplot(.,aes(x=f1_cent, y=vowel, color=language))+
  geom_point()
```

## 3. F2 as a function of vowel and language

```{r}
vowels %>% 
  ggplot(.,aes(x=f2_cent, y=vowel, color=language))+
  geom_point()
```

## 4. (Attempt at) Plotting trajectory length in F1/F2 vowel space

```{r}
bonus_data <- select(my_data,-id, - item, -tl) %>% 
  group_by(vowel, language) %>% 
  summarize(f1_20=mean(f1_20), f1_35=mean(f1_35), f1_50=mean(f1_50),
            f1_65=mean(f1_65), f1_80=mean(f1_80), f2_20=mean(f2_20),
            f2_35=mean(f2_35), f2_50=mean(f2_50),
            f2_65=mean(f2_65), f2_80=mean(f2_80))

bonus<-bonus_data %>%
   pivot_longer(cols = c ("f1_20", "f1_35", "f1_50", "f1_65", "f1_80"), 
                 names_to = "F1", values_to = "value") %>% 
  separate(col=F1, into=c("f1", "time"), sep = "_") %>% 
  pivot_wider(names_from = f1, values_from = value) %>% 
  pivot_longer(cols = c ("f2_20", "f2_35", "f2_50", "f2_65", "f2_80"), 
                 names_to = "F2", values_to = "value") %>% 
  separate(col=F2, into=c("f2", "time"), sep = "_") %>% 
  pivot_wider(names_from = f2, values_from = value)
  
bonus %>%  
  ggplot(., aes(x = f2, y = f1, color=vowel, shape=language)) + 
  geom_point(alpha = 2) +
  geom_text(data = bonus, aes(label = vowel), size = 2) + 
    scale_y_reverse() + 
  scale_x_reverse()
```

I ended up with the plot above; not sure if it works though.
Also, I tried to connect the points in the plot with geom_line and it looks like this (which is kind of difficult to interpret and look at, so I probably missed a step or something):

```{r}
bonus %>%  
  ggplot(., aes(x = f2, y = f1, color=vowel, shape=language)) + 
  geom_point(alpha = 2) + geom_line()+
  geom_text(data = bonus, aes(label = vowel), size = 2) + 
    scale_y_reverse() + 
  scale_x_reverse()
```